MS Depression Paper, started 2022 08 24

### 2022-09-03 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_just_volumes_started_20220824.Rmd
runs all the way through!

``` {r bootstrap}
#### bootstrap

@@

@@

```{r sensitivity analyses}
tertile <- round((dim(fascicle_values_dep_no_cranial_nerves)[1] / 3),0)
quintile <- round((dim(fascicle_values_dep_no_cranial_nerves)[1] / 5),0)

dep_network_names_top_tertile_no_cranial_nerves_by_vol<- (fascicle_values_dep_no_cranial_nerves[order(fascicle_values_dep_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:tertile] #n=26

dep_network_names_top_quintile_no_cranial_nerves_by_vol<- (fascicle_values_dep_no_cranial_nerves[order(fascicle_values_dep_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:quintile] #n=15

dep_network_names_bottom_tertiles_no_cranial_nerves_by_vol <- fascicle_values_dep_no_cranial_nerves$fascicle[!(fascicle_values_dep_no_cranial_nerves$fascicle %in% dep_network_names_top_tertile_no_cranial_nerves_by_vol)]

dep_network_names_bottom_quintiles_no_cranial_nerves_by_vol <- fascicle_values_dep_no_cranial_nerves$fascicle[!(fascicle_values_dep_no_cranial_nerves$fascicle %in% dep_network_names_top_quintile_no_cranial_nerves_by_vol)]

top_tertile_total_volume_no_cranial_nerves <-
sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_tertile_no_cranial_nerves_by_vol)])

top_quintile_total_volume_no_cranial_nerves <-
sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_quintile_no_cranial_nerves_by_vol)])

bottom_tertiles_total_volume_no_cranial_nerves <-
sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_bottom_tertiles_no_cranial_nerves_by_vol)])

bottom_quintiles_total_volume_no_cranial_nerves<-
sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_bottom_quintiles_no_cranial_nerves_by_vol)])

dep_and_healthy_groups_for_ICD_analysis <- dep_and_healthy_groups_for_ICD_analysis %>%

  #tertile and quintiles
  mutate(sum_fascicle_vol_lost_top_tertile = 
           rowSums(across(dep_network_names_top_tertile_no_cranial_nerves_by_vol))) %>%
  mutate(proportion_volume_lost_top_tertile_by_network_size_sum = sum_fascicle_vol_lost_top_tertile /top_tertile_total_volume_no_cranial_nerves) %>%

  mutate(sum_fascicle_vol_lost_bottom_tertiles = 
           rowSums(across(dep_network_names_bottom_tertiles_no_cranial_nerves_by_vol))) %>%
  mutate(proportion_volume_lost_bottom_tertiles_by_network_size_sum = sum_fascicle_vol_lost_bottom_tertiles /bottom_tertiles_total_volume_no_cranial_nerves) %>%

  mutate(sum_fascicle_vol_lost_top_quintile = 
           rowSums(across(dep_network_names_top_quintile_no_cranial_nerves_by_vol))) %>%
  mutate(proportion_volume_lost_top_quintile_by_network_size_sum = sum_fascicle_vol_lost_top_quintile /top_quintile_total_volume_no_cranial_nerves) %>%

  mutate(sum_fascicle_vol_lost_bottom_quintiles = 
           rowSums(across(dep_network_names_bottom_quintiles_no_cranial_nerves_by_vol))) %>%
  mutate(proportion_volume_lost_bottom_quintiles_by_network_size_sum = sum_fascicle_vol_lost_bottom_quintiles /bottom_quintiles_total_volume_no_cranial_nerves) %>%

@@

diagnosis_vector <- c(1,2,1,2) #1 is healthy, 2 is depressed
network_vector <- c(1,1,2,2) #1 is depression network, 2 is non_depression network
mean_diagnosis_by_network_vector <- c(mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]),
                             mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
                             mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]),
                             mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]))

@@

#p=2.2*10^16, top tertile prop vol affected = 0.420, bottom 2 tertiles = 0.298
msdisease_within_versus_outside_depression_network_wilcoxon_tertile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_tertile_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_tertiles_by_network_size_sum,paired = TRUE)

>>>>

@@

##### bootstrap

#### bootstrap

@@

```{r sensitivity analyses}

## to check different splits of the networks

tertile <- round((dim(fascicle_volumes_dep_network_no_cranial_nerves)[1] / 3),0)
quintile <- round((dim(fascicle_volumes_dep_network_no_cranial_nerves)[1] / 5),0)

## get names of each network, and add suffix
dep_network_names_top_tertile_no_cranial_nerves<- (fascicle_volumes_dep_network_no_cranial_nerves[order(fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:tertile]  #n=26

dep_network_names_bottom_tertiles_no_cranial_nerves <- fascicle_volumes_dep_network_no_cranial_nerves$fascicle[!(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_tertile_no_cranial_nerves)]

dep_network_names_top_quintile_no_cranial_nerves<- (fascicle_volumes_dep_network_no_cranial_nerves[order(fascicle_volumes_dep_network_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:quintile]  #n=15

dep_network_names_bottom_quintiles_no_cranial_nerves <- fascicle_volumes_dep_network_no_cranial_nerves$fascicle[!(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_quintile_no_cranial_nerves)]

### get volumes of each network
top_tertile_total_volume_no_cranial_nerves <-
sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_tertile_no_cranial_nerves)])

bottom_tertiles_total_volume_no_cranial_nerves <-
sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_bottom_tertiles_no_cranial_nerves)])

top_quintile_total_volume_no_cranial_nerves <-
sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_top_quintile_no_cranial_nerves)])

bottom_quintiles_total_volume_no_cranial_nerves<-
sum(fascicle_volumes_dep_network_no_cranial_nerves$num_voxels_total[which(fascicle_volumes_dep_network_no_cranial_nerves$fascicle %in% dep_network_names_bottom_quintiles_no_cranial_nerves)])

## change suffix to _vol

dep_network_names_top_tertile_no_cranial_nerves <- dep_network_names_top_tertile_no_cranial_nerves %>% 
  gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE)

dep_network_names_bottom_tertiles_no_cranial_nerves <- dep_network_names_bottom_tertiles_no_cranial_nerves %>% 
  gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE)

dep_network_names_top_quintile_no_cranial_nerves <- dep_network_names_top_quintile_no_cranial_nerves %>%
  gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE)

dep_network_names_bottom_quintiles_no_cranial_nerves <- 
dep_network_names_bottom_quintiles_no_cranial_nerves %>% 
  gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE)

### make new columns
dep_and_healthy_groups_for_ICD_analysis <- dep_and_healthy_groups_for_ICD_analysis %>%

  #tertile and quintiles
  mutate(sum_fascicle_vol_lost_top_tertile = 
           rowSums(across(all_of(dep_network_names_top_tertile_no_cranial_nerves)))) %>%
  mutate(proportion_volume_lost_top_tertile_by_network_size_sum = sum_fascicle_vol_lost_top_tertile /top_tertile_total_volume_no_cranial_nerves) %>%

  mutate(sum_fascicle_vol_lost_bottom_tertiles = 
           rowSums(across(dep_network_names_bottom_tertiles_no_cranial_nerves))) %>%
  mutate(proportion_volume_lost_bottom_tertiles_by_network_size_sum = sum_fascicle_vol_lost_bottom_tertiles /bottom_tertiles_total_volume_no_cranial_nerves) %>%

  mutate(sum_fascicle_vol_lost_top_quintile = 
           rowSums(across(dep_network_names_top_quintile_no_cranial_nerves))) %>%
  mutate(proportion_volume_lost_top_quintile_by_network_size_sum = sum_fascicle_vol_lost_top_quintile /top_quintile_total_volume_no_cranial_nerves) %>%

  mutate(sum_fascicle_vol_lost_bottom_quintiles = 
           rowSums(across(dep_network_names_bottom_quintiles_no_cranial_nerves))) %>%
  mutate(proportion_volume_lost_bottom_quintiles_by_network_size_sum = sum_fascicle_vol_lost_bottom_quintiles /bottom_quintiles_total_volume_no_cranial_nerves)

@@

#diagnosis_vector <- c(1,2,1,2) #1 is healthy, 2 is depressed
#network_vector <- c(1,1,2,2) #1 is depression network, 2 is non_depression network
#mean_diagnosis_by_network_vector <- c(mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]),
#                             mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_dep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
#                             mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]),
 #                            mean(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_nondep_net_by_network_size_sum[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]))

@@

##### within versus outside network
#p=2.2*10^16, top tertile prop vol affected = 0.420, bottom 2 tertiles = 0.298
msdisease_within_versus_outside_depression_network_wilcoxon_tertile <- wilcox.test(dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_top_tertile_by_network_size_sum, dep_and_healthy_groups_for_ICD_analysis$proportion_volume_lost_bottom_tertiles_by_network_size_sum,paired = TRUE)

#Diagnostic effects
